---
title: "size-term-estimation"
author: "suzanne"
date: "2023-04-13"
output: html_document
---

# Read in libraries, set working directory, source code for doing the size estimation work
```{r, setup}
knitr::opts_knit$set(root.dir = "R:/e2projects_two/activitysim/estimation/2017_2019_data/size_terms")
```

```{r}
library(dplyr)
library(tidyverse)
# for mapping places that have people going to them, but size variables are zero:
library(sf)
library(leaflet)
library(htmlwidgets)
library(stats)
library(data.table)

#put your input data here, outputs will also be written out here
setwd("R:/e2projects_two/activitysim/estimation/2017_2019_data/size_terms")

#this file does the work
source('C:/Joanne_PSRC/travel_models/psrc_activitysim/scripts/size_terms/size_estimation_dest_choice.R')

```

There are 5 estimation files for the following destination types:
regular work location 
school location
non-mandatory tour
work-based subtour aka atwork
intermediate stop aka trip destination

```{r}

work_loc_lu<-read.csv('size_term_17_19_workplace.csv') %>% # work trips
  mutate(log_hh_p = log(1+hh_p))
school_loc_lu<-read.csv('size_term_17_19_school.csv') %>%  # school trips
  mutate(log_hh_p = log(1+hh_p))
# non_mandatory_tour_lu<-read.csv('size_term_17_19_non_mandatory.csv') # non-mandatory primary
wb_sub_lu<- read.csv('size_term_17_19_atwork.csv') %>%
  mutate(log_hh_p = log(1+hh_p))
# trip_dest_lu<-read.csv('size_term_17_19_trip.csv')
all_trip_dest_lu<-read.csv('C:/Joanne_PSRC/travel_models/psrc_activitysim/scripts/size_terms/size_term_17_19_all_trip.csv') %>%
  mutate(log_hh_p = log(1+hh_p))

```




write out the list of columns that could be used in the estimation
```{r}

var_list<-c('empedu_p',	# education jobs
            'empfoo_p', # food jobs
            'empgov_p',	# government jobs
            'empind_p',	# industrial job
            'empmed_p',	# medicine jobs?
            'empofc_p',	# office jobs
            'empoth_p',	# other jobs?
            'empret_p',	# retail jobs
            # 'emprsc_p',	# resource jobs
            'empsvc_p',	# service jobs
            'emptot_p',	# total jobs
            # 'hh_p',	    # number of households?
            "log_hh_p",
            'stugrd_p',	
            'stuhgh_p',	
            'stuuni_p',	
            'log_aparks')
all_lu_vars<-data.frame(variable=var_list,Zeros=rep(0,(length(var_list))))
 

```

```{r}
summary(work_loc_lu[,var_list])
```


```{r correlation, fig.width = 9, fig.height = 7}
library(corrplot)
M<-cor(work_loc_lu[,var_list])
head(round(M,2))
corrplot(M, method="number")
```
```{r count sample size}
count_jobs <- work_loc_lu %>%
  mutate(dum_low = ifelse(work_low>0,"low",0),
         dum_med = ifelse(work_med>0,"med",0),
         dum_high = ifelse(work_high>0,"high",0),
         dum_very_high = ifelse(work_very_high>0,"very_high",0))

# count zones that have workers in each income bin
fn <- function(.data,var){
  .data %>% 
    add_row(count_jobs %>%
    group_by({{var}}) %>%
    summarise_at(vars(empedu_p:emptot_p),~sum(.>0)) %>%
    rename(worker_inc = {{var}}) %>%
    ungroup())
}
# sum all jobs in zones that have workers in each income bin
test <-count_jobs %>%
  group_by(dum_low) %>%
  summarise_at(vars(empedu_p:emptot_p),~sum(.>0)) %>%
  rename(worker_inc = dum_low) %>%
  fn(dum_med) %>%
  fn(dum_high) %>%
  fn(dum_very_high) %>%
  filter(worker_inc!="0")
write.csv(test,"outputs/mle_sample_size_workplace.csv")
getwd()
```

all vars: empsvc_p, empedu_p,	empfoo_p, empgov_p,	empind_p,	empmed_p,	empofc_p,	empoth_p,	empret_p,	hh_p, log_aparks
## Regular Work Location
```{r workplace_work_low}
workplace_work_low <- estimate_size_terms(work_loc_lu, 
                    model = "workplace", 
                    seg = "work_low", 
                    select_var_list=c("work_low",	
                                      'empsvc_p',	
                                      'empfoo_p',
                                      'empedu_p', 
                                      'empgov_p',	
                                      'empind_p',	
                                      'empmed_p',	
                                      'empofc_p',	
                                      # 'empoth_p',	
                                      'empret_p'),# all variables
                    all_lu_vars=all_lu_vars)

```


```{r workplace_work_med}
workplace_work_med<-estimate_size_terms(work_loc_lu, 
                    model = "workplace", 
                    seg = "work_med", 
                    select_var_list=c("work_med",
                                      'empsvc_p', 
                                      'empfoo_p', # food jobs
                                      'empedu_p',	# education jobs
                                      'empgov_p',	# government jobs
                                      'empind_p',	# industrial job
                                      'empmed_p',	# medicine jobs?
                                      'empofc_p',	# office jobs
                                      # 'empoth_p',	# other jobs?
                                      'empret_p'),
                    all_lu_vars=all_lu_vars)
```

```{r workplace_work_high}
workplace_work_high<-estimate_size_terms(work_loc_lu,  
                    model = "workplace", 
                    seg = "work_high", 
                    select_var_list=c("work_high",  
                                      'empofc_p',
                                      'empsvc_p',	
                                      'empfoo_p', 
                                      'empedu_p',	
                                      'empgov_p',	
                                      'empind_p',	
                                      'empmed_p',	
                                      # 'empoth_p',	
                                      'empret_p'),
                    all_lu_vars=all_lu_vars)
```
```{r workplace_work_very_high}
workplace_work_very_high <- estimate_size_terms(work_loc_lu,  
                    model = "workplace", 
                    seg = "work_very_high", 
                    select_var_list=c("work_very_high", 
                                      'empofc_p',	# office jobs
                                      'empsvc_p', 
                                      'empfoo_p', # food jobs
                                      'empedu_p',	# education jobs
                                      'empgov_p',	# government jobs
                                      'empind_p',	# industrial job
                                      'empmed_p',	# medicine jobs?
                                      'empoth_p',	# other jobs?
                                      'empret_p'),
                    all_lu_vars=all_lu_vars)
```
# University can't get this model to estimate when include stuuni_p

```{r}
school_loc_lu_uni<- school_loc_lu%>%
select(zone, university,   stuuni_p, empedu_p)
```


won't estimate

```{r}
#estimate_size_terms(school_loc_lu_uni, 'outputs/school_university.csv', all_lu_vars)
```

```{r}
ggplot(school_loc_lu_uni, aes(x=stuuni_p, y=university))+geom_point()

```





```{r}
ggplot(school_loc_lu_uni, aes(x=empedu_p, y=university))+geom_point()+geom_smooth( method="lm", se = TRUE)

```
Alternatively estimate a linear model and see what happens
```{r}
fitted_model_university<-glm(university ~ stuuni_p+ empedu_p-1, data=school_loc_lu_uni)
summary(fitted_model_university)
```


## Gradeschool
# won't estimate

```{r}
school_loc_lu_gradeschool<- school_loc_lu%>%
select(zone, gradeschool, stugrd_p,empedu_p)
```

## won't estimate
```{r}
#estimate_size_terms(school_loc_lu_gradeschool, 'outputs/school_gradeschool.csv', all_lu_vars)
```


```{r}
fitted_model_gradeschool<-glm(gradeschool ~ stugrd_p +empedu_p-1, data=school_loc_lu_gradeschool)
summary(fitted_model_gradeschool)
```
## high school
```{r}
school_loc_lu_highschool<- school_loc_lu%>%
select(zone, highschool, stuhgh_p,empedu_p)
```

## won't estimate
```{r}
#estimate_size_terms(school_loc_lu_highschool, 'outputs/school_highschool.csv', all_lu_vars)
```

```{r}
fitted_model_highschool<-glm(highschool ~ stuhgh_p +empedu_p-1, data=school_loc_lu_highschool)
summary(fitted_model_highschool)
```

## Non-mandatory primary tour destinations and intermediate trip destinations - escort

```{r work_size}
non_mandatory_work<-estimate_size_terms(all_trip_dest_lu, 
                    model = "non_mandatory", 
                    seg = "work",  
                    select_var_list=c("work", 
                                      "empofc_p","empsvc_p", "empedu_p",	"empfoo_p", "empgov_p"),
                    all_lu_vars=all_lu_vars)
```
```{r univ}
non_mandatory_univ<-estimate_size_terms(all_trip_dest_lu, 
                    model = "non_mandatory", 
                    seg = "univ",  
                    select_var_list=c("univ",
                                      "stuuni_p", # TODO: do we want to have both employment and enrollment in here?
                                      "empedu_p"),
                    all_lu_vars=all_lu_vars)
```

```{r non_mandatory_escort}
non_mandatory_escort <- estimate_size_terms(all_trip_dest_lu, 
                    model = "non_mandatory", 
                    seg = "escort", 
                    select_var_list=c("escort", 
                                      'empsvc_p',	# service jobs
                                      'empfoo_p',
                                      'empret_p',	# retail jobs
                                      'hh_p',	    # number of households?
                                      'stugrd_p',	
                                      'stuhgh_p'
                                      ),
                    all_lu_vars=all_lu_vars)
```

```{r non_mandatory_shopping}
non_mandatory_shopping<- estimate_size_terms(all_trip_dest_lu, 
                    model = "non_mandatory", 
                    seg = "shopping", 
                    select_var_list=c("shopping",
                                      'empret_p', 
                                      'empsvc_p',	# service jobs
                                      'empfoo_p'
                                      ),
                    all_lu_vars=all_lu_vars)
```

## Non-mandatory eatout
```{r non_mandatory_eat}
non_mandatory_eatout<-estimate_size_terms(all_trip_dest_lu, 
                    model = "non_mandatory", 
                    seg = "eatout", 
                    select_var_list=c("eatout", 
                                      'empfoo_p',
                                      'empret_p'
                                      ),
                    all_lu_vars=all_lu_vars)
```
```{r non_mandatory_othmaint}
non_mandatory_othmaint<-estimate_size_terms(all_trip_dest_lu, 
                    model = "non_mandatory", 
                    seg = "othmaint", 
                    select_var_list=c("othmaint", 
                                      "empsvc_p",
                                      'empfoo_p',"empmed_p",
                                      #"empgov_p",	"empofc_p",	
                                      "empret_p"),
                    all_lu_vars=all_lu_vars)
```

```{r non_mandatory_social}
non_mandatory_social<-estimate_size_terms(all_trip_dest_lu, 
                    model = "non_mandatory", 
                    seg = "social", 
                    select_var_list=c("social", 
                                      "log_hh_p",
                                      "empsvc_p", "empfoo_p",	
                                      "empret_p", # TODO: keep or not keep?
                                      "log_aparks"),
                    all_lu_vars=all_lu_vars)

```
```{r non_mandatory_othdiscr}

non_mandatory_othdiscr<-estimate_size_terms(all_trip_dest_lu, 
                    model = "non_mandatory", 
                    seg = "social",
                    select_var_list=c("othdiscr", 
                                      "empsvc_p", "empedu_p",	"empfoo_p","empret_p",	"log_hh_p",
                                      "log_aparks"),
                    all_lu_vars=all_lu_vars)
```
## At Work

wb_sub_lu

empsvc_p, empedu_p,	empfoo_p, empgov_p,	empind_p,	empmed_p,	empofc_p,	empoth_p,	empret_p,	hh_p, log_aparks
```{r atwork}
atwork<- estimate_size_terms(wb_sub_lu, 
                    model = "atwork", 
                    seg = "atwork",
                    select_var_list=c("atwork", 
                                      "empfoo_p","empsvc_p","empofc_p","empmed_p","empret_p",
                                      "log_aparks"),
                    all_lu_vars=all_lu_vars)
```

```{r save all size terms}
size_terms_no_school <- rbind(workplace_work_low,
                    workplace_work_med,
                    workplace_work_high,
                    workplace_work_very_high,
                    non_mandatory_work,
                    non_mandatory_univ,
                    non_mandatory_escort,
                    non_mandatory_shopping,
                    non_mandatory_eatout,
                    non_mandatory_othmaint,
                    non_mandatory_social,
                    non_mandatory_othdiscr,
                    atwork)

create_school_size <- function(model,seg){
  df <- data.frame(matrix(c(model,seg,integer(15)), nrow=1))
  colnames(df) <- names(size_terms_no_school)
  return(df)
}
school_gradeschool <- create_school_size("school","gradeschool") %>% 
  mutate(stugrd_p=1)
school_highschool <- create_school_size("school","highschool") %>% 
  mutate(stuhgh_p=1)
school_university <- create_school_size("school","university") %>% 
  mutate(stuuni_p=1)

size_terms <- rbind(size_terms_no_school,
                    school_gradeschool,
                    school_highschool,
                    school_university
                    )
write.csv(size_terms,
          "outputs/destination_choice_size_terms_psrc/size_terms_mle_all_trips_all_variables_for_workplace.csv",
          row.names = FALSE)
write.csv(size_terms %>% filter(model_selector!="workplace"),
          "outputs/destination_choice_size_terms_psrc/size_terms_mle_all_trips_no_workplace.csv",
          row.names = FALSE)

```

```{r pums work locations and other size terms}

```

### Intermediate Stops/Trip Ends

```{r work_size}
estimate_size_terms(trip_dest_lu, 
                    outfilename='outputs/work_size.csv', 
                    select_var_list=c("work", 
                                      "empofc_p","empsvc_p", "empedu_p",	"empfoo_p", "empgov_p"),
                    all_lu_vars=all_lu_vars)
```
```{r}

estimate_size_terms(trip_dest_lu, 
                    outfilename='outputs/escort_size.csv', 
                    select_var_list=c("escort", 
                                      "empsvc_p","empedu_p",  "empret_p", "empmed_p", "stugrd_p",
                                      "stuhgh_p", "log_aparks"),
                    all_lu_vars=all_lu_vars)

```
```{r shopping_size}
estimate_size_terms(trip_dest_lu, 
                    outfilename='outputs/shopping_size.csv', 
                    select_var_list=c("shopping", 
                                      "empret_p", "empfoo_p", "empsvc_p"),
                    all_lu_vars=all_lu_vars)

```

```{r eatout_size}
estimate_size_terms(trip_dest_lu, 
                    outfilename='outputs/eatout_size.csv', 
                    select_var_list=c("eatout", 
                                      "empfoo_p", "empret_p", "empsvc_p"),
                    all_lu_vars=all_lu_vars)

```


```{r othmaint_size}
estimate_size_terms(trip_dest_lu, 
                    outfilename='outputs/othmaint_size.csv', 
                    select_var_list=c("othmaint", 
                                      "empfoo_p","empsvc_p","empgov_p", "empmed_p","empret_p",
                                      "log_aparks"),
                    all_lu_vars=all_lu_vars)
```


```{r social_size}
estimate_size_terms(trip_dest_lu, 
                    outfilename='outputs/social_size.csv', 
                    select_var_list=c("social", 
                                      "empsvc_p", "empfoo_p", "hh_p", "empret_p"),
                    all_lu_vars=all_lu_vars)

```



```{r other_discr_size}
estimate_size_terms(trip_dest_lu, 
                    outfilename='outputs/other_discr_size.csv', 
                    select_var_list=c("othdiscr", 
                                      "empsvc_p","empedu_p","empfoo_p","empret_p","hh_p",
                                      "log_aparks"),
                    all_lu_vars=all_lu_vars)
```

# University
I'm going to constrain this to one. I'm just not sure we want university students attracted to education jobs.
```{r univeristy_size}
estimate_size_terms(trip_dest_lu, 
                    outfilename='outputs/univeristy_size.csv', 
                    select_var_list=c("univ", 
                                      "empedu_p","stuuni_p"),
                    all_lu_vars=all_lu_vars)

```




